---
title: "Dplyr Notes"
author: "carlhank"
output: html_document
runtime: shiny
---

### dplyr main verb(function)

+ filter    — 選取一個資料框架列(row)的子集。
+ arrange   — 對資料框架中變量的內容進行重新排序。
+ select    — 選取需要資料框架中所需的變量(columns)。
+ mutate    — 添加變數(columns)到已存在的資料框架中。
+ summarize — summarize values.
+ group_by  — 描述如何將資料框架分割成幾個群組。

※ chain operator：`%.%`, 結合(combine)上述函數以快速方便操作資料。


### Install and library
Install：
```r
install.packages("dplyr")
```

Library：
```{r echo = TRUE}
library(dplyr)
```


### ToothGrowth
Here we will use a data set that comes with R called `ToothGrowth`.

```{r, echo = FALSE}
numericInput("rows", "Control number of rows of all tables", value = 6, min = 0, max = dim(ToothGrowth)[1])
```

```r
head(ToothGrowth)
```
```{r, echo = FALSE}
renderTable({
    head(ToothGrowth, n = input$rows)
})
```
```{r, echo = TRUE}
str(ToothGrowth)
```

+ Data description
    - len: the length of teeth in each of 10 guinea pigs
    - supp: two delivery methods (orange juice or ascorbic acid)
    - dose: 維他命C劑量濃度(0.5, 1 and 2 mg)


### Basic usage

**1.** `filter()`
```r
filter(ToothGrowth, supp == "VC", dose == 1.0), n = input$rows
```
```{r, echo = FALSE}
renderTable({
    head(filter(ToothGrowth, supp == "VC", dose == 1.0), n = input$rows)
})
```
This is equivalent to R basic indicator method:
```r
ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0, ]
```
```{r, echo = FALSE}
renderTable({
    head(ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0, ], n = input$rows)
})
```
Or you can use `subset` function:
```r
subset(ToothGrowth, supp == "VC" & dose == 1.0, drop = TRUE)
```
```{r, echo = FALSE}
renderTable({
    head(subset(ToothGrowth, supp == "VC" & dose == 1.0, drop = TRUE), n = input$rows)  
})
```


**2.** `arrange()`
EX1. `ToothGrowth` data
```r
arrange(ToothGrowth, len)
arrange(ToothGrowth, desc(len))
```
```{r, echo = FALSE}
renderTable({
    head(arrange(ToothGrowth, len), n = input$rows)
    head(arrange(ToothGrowth, desc(len)), n = input$rows)
})
```
EX2. `iris` data
```r
head(iris)
arrange(iris, Sepal.Length, desc(Sepal.Width))  ## increase Sepal.Length and descrease Sepal.Width
```
```{r, echo = FALSE}
renderTable({
    head(iris, n = input$rows)
    head(arrange(iris, Sepal.Length, desc(Sepal.Width)), n = input$rows)
})
```


**3.**`select()`
EX1. `ToothGrowth` data
```r
select(ToothGrowth, len, supp)
select(ToothGrowth, supp, dose)
```
```{r, echo = FALSE}
renderTable({
    head(select(ToothGrowth, len, supp), n = input$rows)
    head(select(ToothGrowth, supp, dose), n = input$rows)
})
```
EX2. `iris` data
```r
select(iris, Sepal.Width, Petal.Width, Species)
select(iris, -c(Sepal.Length, Petal.Length))
```
```{r, echo = FALSE}
renderTable({
    head(select(iris, Sepal.Width, Petal.Width, Species), n = input$rows)
    head(select(iris, -c(Sepal.Length, Petal.Length)), n = input$rows)
})
```


**4.** `mutate()`
```r
mutate(ToothGrowth, meanLen = mean(len), sdLen = sd(len))
```
```{r, echo = FALSE}
renderTable({
    head(mutate(ToothGrowth, meanLen = mean(len), sdLen = sd(len)),
         n = input$rows)
})
```
This is equivalent to R basic `transform` function:
```r
transform(ToothGrowth, meanLen = mean(len), sdLen = sd(len))
```
```{r, echo = FALSE}
renderTable({
    head(transform(ToothGrowth, meanLen = mean(len), sdLen = sd(len)),
         n = input$rows)
})
```


**5.** `summarize()` or `summarise()`
```{r, echo = TRUE}
summarize(ToothGrowth, meanLen = mean(len))
summarise(ToothGrowth, sdLen = sd(len))
```


### Chain operator `%.%` usage

**1.** Calculate the mean length of teeth (len) for each delivery method (supp)
```r
ToothGrowth %.%                      ## data
group_by(supp) %.%                   ## group
summarize(meanLen = mean(len))       ## calculate mean
```
```{r, echo = FALSE}
renderTable({
    head(ToothGrowth %.%                      ## data
         group_by(supp) %.%                   ## group
         summarize(meanLen = mean(len))       ## calculate mean)
         , n = input$rows)
})
```


**2.** Calculate the mean teeth length by delivery method and dose level?
```{r, echo = TRUE}
ToothGrowth %.%
group_by(supp, dose) %.%
summarise(meanLen = mean(len), sdLen = sd(len))
```


#####*Now please compare the following 3. and 4. results：*

**3.** Create new variable
```r
ToothGrowth %.%
mutate(StdLen = (len - mean(len)) / sd(len))
```
```{r, echo = FALSE}
renderTable({
    head(ToothGrowth %.%
         mutate(StdLen = (len - mean(len)) / sd(len)),
         n = input$rows)
})
```


**4.** Create new variable, which is the standard value of the tooth length in the six supp x dose groups.
```r
ToothGrowth %.%
group_by(supp, dose) %.%
mutate(StdLen = (len - mean(len)) / sd(len))
```
```{r, echo = FALSE}
renderTable({
    head(ToothGrowth %.%
         group_by(supp, dose) %.%
         mutate(StdLen = (len - mean(len)) / sd(len)),
         n = input$rows)
})
```

##### Practice

**[practice 1]**
```{r, echo = TRUE}
ToothGrowth %.%
filter(dose == 1) %.%
select(len, supp) %.%
arrange(len)
```


**[practice 2]**
```{r, echo = TRUE}
TG05 <- ToothGrowth %.%
filter(dose == 0.5) %.%
select(len, supp) %.%
arrange(desc(len)) %.%
head(n = 5)
```
```{r, echo = TRUE}
TG05
```

